%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%                                                                 %
%                            ABSTRACT                             %
%                                                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 
\specialhead{ABSTRACT}
 
Provenance information is critical for research publication readers to correctly interpret report content and enables them to evaluate the report by digging into the software in use, source data and responsible agents. However, creating proper provenance information for research publications is very hard. First, it requires knowledge of proper provenance information to capture for the report creating process, causing either extra learning overhead on the domain scientists who write the reports or communication overhead between domain scientists and informatics experts who keep track of the provenance information. Second, it requires extra efforts on top of the authoring work, which already includes writing manuscripts, getting data from various sources, and running programs to generate tables and figures, so the creation of provenance information is usually insufficiently motivated. Third, since provenance information is usually closely related to the running environment of the programs used during the process of preparing publications, it is often necessary to also capture the configurations of program running platform such as the operating system or even the computer hardware in order to get useful provenance information for the purpose of enabling publication readers to reproduce and validate the content, which requires knowledge of the computational system. This knowledge is usually outside the reach of both domain scientists and informatics experts, so they would depend on system administrators to provide them with or learn by themselves the configuration related provenance information, causing more communication burden and complexity in provenance capturing.

Existing scientific workflow management systems such as VisTrails \cite{freire2014reproducibility}, Kepler \cite{ludascher2006scientific}, Taverna \cite{wolstencroft2013taverna} and ReproZip \cite{chirigati2013reprozip} do not provide satisfactory solutions to any of the three problems mentioned above. Users need to learn how to use these systems in order to create provenance information with them. These systems require their users to take the extra efforts of integrating the provenance information into the workflow systems, and usually there is not enough incentive to do this job after writing the publications. Information for program running platforms either is not taken care of at all, or heavily depends on the specific system architecture, e.g., by tracking operating system calls, as in the case of the ReproZip system. \comment{(more related work to add...)}

In this proposal, I propose a paradigm of preparing research publications to overcome all the problems associated with provenance capturing in the preparing process, which is to create publications with libraries that transparently capture provenance information on a portable platform.

Transparent provenance capturing means the users of the libraries are not exposed to the provenance information they are capturing, which is captured at the time of each publication preparing action being taken, e.g., when a block of manuscript is written, when some data are downloaded from an external source, or when a plot is created by calling a library function with the downloaded data. Capturing provenance information in this way eliminates the need to ask the authors to explicitly input provenance information, thus it also removes from the authors the burden of learning what provenance information to capture or communicating with people who are familiar with provenance capturing. The assumption is that the libraries can be preloaded with proper ontologies for domain specific provenance capturing.

A portable platform is a software package or a Web application that runs on some environment that is available on all the popular computational systems, such as a Java program which runs on the Java Virtual Machine, a Python program which runs with the Python interpreter, or a Web based application written in HTML, JavaScript and/or PHP scripts which runs on any of the mainstream Web browsers. Preparing research publications based entirely on a portable platform eliminates the necessity of capturing provenance information at the computer system level. Two things desirable for the platform are that 1) it should supply all the functions needed in creating research publications, and 2) these functions can be used in a way that is similar to the software packages the writer is used to working with. To ensure the first point, the platform needs to be extensible, better if it already has a strong development community to supply a plethora of evolving modules. The second point requires that the platform supports popular interactive editing modes widely accepted in the research community; examples include those seen in research software packages such as MATLAB and R.

Contributions of the proposed work are 1) the creation of a framework that evaluates provenance capturing systems for research publications in terms of tool learning and usage, explicit information input and system level provenance capturing necessity, 2) the proposal of a provenance capturing paradigm that is optimal according to the framework, and 3) the implementation of a provenance aware publication editing library for a specific research domain that supports a paradigm that is close to the optimal. \comment{(I'm still not able to put them nicely...)}

\comment{A library that transparently captures provenance information enables researchers to focus on the science part of creating publications, without being distracted by the provenance preserving part. As long as the functions provided by the library are invoked how to ensure functionalities active development community}

\comment{Typical activities include importing a library to the workspace, calling a function to get source data or to create an image, adding a text block to the publication, and rerunning a code block with a different set of parameters.}

\comment{The challenges, existing efforts and their shortcomings will be written here.}

\comment{Feasibility comes here.}

\comment{Contributions come here.}
